Analyzing Arrest Patterns and Racial Representation in Toronto

### 2024-12-03 paper/paper.qmd
Adding pdf of paper and changed beta

---
title: "Analyzing Arrest Patterns and Racial Representation in Toronto"
author: "Akshat Aneja"
abstract: This paper analyzes arrest data from the Toronto Police Service for the years 2020 and 2021, focusing on the distribution of arrests by ethnic identity, age, and booking status to assess whether systemic bias exists in policing practices. The analysis reveals that, while arrests appear to be driven by objective factors, disparities related to socio-economic status may still exist. These findings provide important insights that can help improve transparency and accountability in policing, contributing to more equitable law enforcement practices.
date: today
date-format: long
number-sections: true
format: pdf
toc: true
thanks: "Code and data supporting this analysis is available at: https://github.com/Akshat211202/Analysis-of-bias-in-arrests/"
bibliography: references.bib
---

@@

\begin{align}
y_i|\pi_i &\sim \mbox{Bern}(\pi_i) \label{eq1} \\
\mbox{logit}(\pi_i) &= \beta_0 + \beta_1 \times \mbox{race}_i + \beta_2 \times \mbox{year}_i \label{eq2} \\
\beta_0 &\sim \mbox{Normal}(0, 5) \label{eq3} \\
\beta_1 &\sim \mbox{Normal}(0, 5) \label{eq4} \\
\beta_2 &\sim \mbox{Normal}(0, 5) \label{eq5}
\end{align}

The Bayesian logistic regression model is represented by Equations (\ref{eq1}), (\ref{eq2}), (\ref{eq3}), (\ref{eq4}), and (\ref{eq5}). In these equations: - $y_i$ represents the booking status of the (i)-th individual (0 for not booked and 1 for booked). - $\pi_i$ represents the probability of the (i)-th individual being booked after arrest. - (\beta\_0) represents the intercept, or the log-odds of being booked for a White individual in the year 2020 (reference group). - (\beta\_1) represents how much the log-odds of being booked change on average depending on the race of the individual (relative to a White individual). - (\text{race}\_i) represents the race of the (i)-th individual (categorical predictor with levels: White, Black, Indigenous, South Asian, East/Southeast Asian, Middle Eastern, Latino, and Unknown/Legacy). - (\beta\_2) represents how much the log-odds of being booked change on average depending on the year (relative to 2020). - (\text{year}\_i) represents the year of the (i)-th individual’s arrest (binary: 0 for 2020, 1 for 2021).

The priors for (\beta\_0), (\beta\_1), and (\beta\_2) are weakly informative normal distributions with a mean of 0 and a standard deviation of 5. These priors allow for flexibility in parameter estimation while regularizing extreme values.

The (\beta) coefficients are important because they provide insight into how an individual’s race and the year of arrest influence the likelihood of being booked after arrest.

>>>>

---
title: "Analyzing Arrest Patterns and Racial Representation in Toronto"
author: "Akshat Aneja"
abstract: This paper analyzes arrest data from the Toronto Police Service for the years 2020 and 2021, focusing on the distribution of arrests by ethnic identity, age, and booking status to assess whether systemic bias exists in policing practices. The analysis reveals that, while arrests appear to be driven by objective factors, disparities related to socio-economic status may still exist. These findings provide important insights that can help improve transparency and accountability in policing, contributing to more equitable law enforcement practices.
date: today
date-format: long
number-sections: true
format: pdf
toc: true
thanks: "Code and data supporting this analysis is available at: https://github.com/Akshat211202/bias-in-arrests-toronto"

bibliography: references.bib
---

@@

\begin{align}
y_i|\pi_i &\sim \mbox{Bern}(\pi_i) \label{eq1} \\
\mbox{logit}(\pi_i) &= \beta_0 + \beta_1 \times \mbox{race}_i + \beta_2 \times \mbox{year}_i \label{eq2} \\
\beta_0 &\sim \mbox{Normal}(0, 2.5) \label{eq3} \\
\beta_1 &\sim \mbox{Normal}(0, 2.5) \label{eq4} \\
\beta_2 &\sim \mbox{Normal}(0, 2.5) \label{eq5}
\end{align}

The Bayesian logistic regression model is represented by Equations (\ref{eq1}), (\ref{eq2}), (\ref{eq3}), (\ref{eq4}), and (\ref{eq5}). In these equations: $y_i$ represents the booking status of the $i^{th}$ individual (0 for not booked and 1 for booked). $\pi_i$ represents the probability of the $i^{th}$ individual being booked after arrest. $\beta_0$ represents the intercept, or the log-odds of being booked for a White individual in the year 2020 (reference group). $\beta_1$ represents how much the log-odds of being booked change on average depending on the race of the individual (relative to a White individual). $\text{race}_i$ represents the race of the $i^{th}$ individual (categorical predictor with levels: White, Black, Indigenous, South Asian, East/Southeast Asian, Middle Eastern, Latino, and Unknown/Legacy). $\beta_2$ represents how much the log-odds of being booked change on average depending on the year (relative to 2020). $\text{year}_i$ represents the year of the $i^{th}$ individual’s arrest (binary: 0 for 2020, 1 for 2021).

The priors for $\beta_0$, $\beta_1$, and $\beta_2$ are weakly informative normal distributions with a mean of 0 and a standard deviation of 5. These priors allow for flexibility in parameter estimation while regularizing extreme values.

The $\beta$ coefficients are important because they provide insight into how an individual’s race and the year of arrest influence the likelihood of being booked after arrest.
